---
title: "LAB 5"
author: "Hassan Mazari"
date: "2025-04-02"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#TASK 1 GETTING STARTED

```{r}
print("Hello, R!")
sessionInfo()
```

# Load libraries
library(readxl)
library(dplyr)
library(ggplot2)
library(caret)
```

# TASK 2 Working with Data Imports

```{r}
data_csv <- read.csv("F:/R_codes/csv_data_iris.csv", header = TRUE, stringsAsFactors = FALSE)
head(data_csv)  # Display the first few rows of the dataset

library(readxl)
iris_data <- read_excel("F:/R_codes/excel_data_iris.xlsx")
head(iris_data)
```

# TASK 3 DATA PREPROCESSING AND SUMMARY

```{r}
clean_iris <- na.omit(iris)  # Remove rows with missing values\

library(dplyr)
filtered_iris <- iris %>% filter(Sepal.Length > 5.5)

summary_data <- iris %>%
    group_by(Species) %>%
    summarize(mean_sepal_length = mean(Sepal.Length))
print(summary_data)
```

# TASK 4 DATA VISUALIZATION

```{r}
#BAR CHART
library(ggplot2)

ggplot(iris, aes(x = Species, y = Sepal.Length)) +
    geom_bar(stat = "summary", fun = "mean") +
    labs(title = "Average Sepal Length per Species")

#SCATTER PLOT
ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
    geom_point() +
    labs(title = "Scatter Plot of Sepal vs Petal Length")

#INTERACTIVE VISUALIZATION
library(plotly)

plot_ly(data = iris, x = ~Sepal.Length, y = ~Petal.Length, 
        type = "scatter", mode = "markers", color = ~Species)
```

TASK 5 ADVANCE ANALYSIS
```{r}
#LINEAR REGRESSION
model <- lm(Petal.Length ~ Sepal.Length, data = iris)
summary(model)  # Check model details

#CLUSTERING
clusters <- kmeans(iris[, 1:4], centers = 3)
iris$Cluster <- as.factor(clusters$cluster)

ggplot(iris, aes(Sepal.Length, Petal.Length, color = Cluster)) +
    geom_point()

#MACINE LEARNING WITH CARET
library(caret)

model <- train(Species ~ ., data = iris, method = "rf", trControl = trainControl(method = "cv", number = 5))
print(model)
```

Save Cleaned Data
```{r}
write.csv(iris, "cleaned_iris.csv", row.names = FALSE)
```

Conclusion

This report analyzed the Iris dataset, performed exploratory data analysis, visualized relationships, and implemented a Random Forest classification model.

